Subjective status, Social Class and meritocracy perception

### 2023-11-29 paper-blind.Rmd
ajustes apa y models

---
 title: "Subjective status, Social Class and meritocracy perception"
 # author: Julio Iturra ^[Bremen International Graduate School of Social Sciences (jiiturra@uni-bremen.de)]
 bibliography: "input/bib/ess-merit.bib"
 csl: "input/bib/apa6.csl"
 css: "input/css/custom.css"
 output:
     bookdown::html_document2:
          number_sections: false
     bookdown::pdf_document2:
         template: default
         toc: false
         number_sections: false
     bookdown::word_document2:
          toc: true
 geometry: margin=1in
 header-includes:
   # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
   - \usepackage{times}           # Times New Roman
   - \usepackage{caption}
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
 linestretch: '1.5'
 fontsize: 12pt
 link-citations: yes
 linkcolor: blue
 editor_options:
   chunk_output_type: inline
 abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
---

@@

<!-- Merit is understood as the combination of effort and ability [@Sen2000]. Thus, the meritocratic ideal is focused on the individual, rather than structural factors in getting ahead in life [@mccall_exposure_2017].  In this regard, the measurement of meritocracy-related constructs has not been clear in distinguishing between descriptive and normative dimensions [@Castilloetal2019]. For instance, it is often to notice undifferentiated use of *Meritocratic Beliefs* [@mijs_how_2021a; @bernardo_meritocracy_2021] and *Perceived Meritocracy* [@Xian2017]. The lack of distinction can be problematic, especially when scrutinizing explanations on subjective views about meritocracy.   -->

<!-- The concept of  _perception_ is understood as the subjective description of an object [@Aalberg2003]. In this regard, the descriptive (*how is*) dimension does not consider the normative evaluation of the object (*should be*). On the other hand, the normative dimension can be understood as preferences, influenced by previous beliefs and information affecting the evaluation of an object [@druckman_preference_2000]. In this regard, @Garcia2001 suggests perceived meritocracy is the degree to which individuals consider their society complies with the meritocratic principles. In other words, to what extent effort and talent are rewarded, indicating "how meritocracy is" rather than "how it should be".  -->

@@

The objective position is measured through three variables. First, _Household Income Decile_ is calculated using monthly per capita household income and categorised into deciles. Second, _Educational level_ corresponds to the five categories for educational attainment following the International Standard Classification of Education [@unesco_international_2012]. Finally, we use an adapted five-category version of the EGP class scheme [@Connelly2016], a commonly used measurement in social stratification research [@solis_class_2019]. The Stata package `isko` was used to compute the class scheme [@hendrickx_isko_2002].

Control variables are age, gender, political identification and the perceived salary gap.

```{r ,include=FALSE}
 # Funciones para renderizar tablas tanto en HTML como en pdf
 if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
 if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
 pacman::p_load(knitr, summarytools, kableExtra, dplyr, lavaan,
                haven, htmlwidgets, webshot, sjPlot,sjlabelled) # librerias
 knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                       message = FALSE,  # mensajes/avisos de librerias
                       cache = FALSE,    # cache de los chunks,usar analisis pesados
                       out.width = '85%',# largo de imagen en %
                       fig.pos= "H",     # posicion figuras H = HERE
                       echo = FALSE      # incluir chunk en output
 )
 # Otras configuraciones
 options(scipen=999) # notacion cientifica
 options(knitr.kable.NA = '') # NA en kable = ''
 # Funciones para renderizar tablas tanto en HTML como en pdf
 fwidth = if(is_html_output()) { #Usar en argumento "full_width" de kableExtra
   T
 } else if(is_latex_output()) {
   F
 }
 graph.magnif_table = if(is_html_output()) { # Usar en argumento "graph.magnif=" de summarytools
   1.2
 } else if(is_latex_output()) {
   0.75
 }
 width_table = if(is_html_output()) {
   800
 } else if(is_latex_output()) { # Usar en "width=" dentro del argumento "image=" en kable
   500
 }
 # Css
 st_css()
 # Cargar base de datos procesada
  load(here::here("input/data/proc/elsoc_w03.RData"))
  elsoc_w03$meffort <-
   factor(elsoc_w03$meffort,
          levels = 1:5,
          labels = c("Strongly disagree","Disagree",
                     "Neither agree nor disagree", "Agree", "Strongly agree"))

  elsoc_w03$egpisko05 <-
   factor(elsoc_w03$egpisko05,levels = levels(elsoc_w03$egpisko05),
          labels = c("Unskilled worker (VII.ab+IVc)",
            "Skilled Manual Worker (V+VI)",
            "Self-Employment (IV.ab)",
            "Routine Non-manual (III)",
            "Service (I+II)"))
  elsoc_w03$egpisko05 <- sjlabelled::set_label(elsoc_w03$egpisko05,"Social Class")

  elsoc_w03$merit <- sjlabelled::set_label(elsoc_w03$merit,"Perceivied meritocracy")

  df_desc<- elsoc_w03 %>%
    dplyr::select(merit,meffort,mtalent,
           ess,inc10h.imp,edcine2,egpisko05,lngap_perc,
           sexo,edad,ppolcat) %>% 
    dplyr::select(-meffort,-mtalent,
                  # -sexo,-edad,
                  # -ppolcat,-lngap_perc
                  )
  varlabs <- sjlabelled::get_label(df_desc) #recuperar labels
  df_desc <- na.omit(df_desc)
  set_label(df_desc) <- varlabs  
 st_options(lang = "en",
            footnote = paste0("Source: Authors calculation based on ELSOC 2018 (N = ",
            dim(df_desc)[[1]],")"),
            bootstrap.css = F,
            # custom.css = "input/css/dfsummary.css"
            )
 df<- dfSummary(df_desc,
                plain.ascii = FALSE,
                style = "grid",
                tmp.img.dir = "/tmp",
                graph.magnif = TRUE,
                headings = F,  # encabezado
                varnumbers = F, # num variable
                labels.col = T, # etiquetas
                na.col = F,    # missing
                graph.col = T, # plot
                valid.col = T,max.string.width = 50,
                max.distinct.values = 8 # n valido 
                # col.widths = c(1000,10,10,10)
                )
 df$Variable <- NULL # delete variable column
 df$Valid <- NULL
 summarytools::view(df,file = "output/tables/desc01.html") # Ver tabla en un archivo HTML
 webshot(url ="output/tables/desc01.html",
         file ="output/tables/desc01.png",
         vwidth = 700,vheight = 600) # Remover archivo HTML
```

@@

```{r plot-likert, fig.height=3, fig.width=10, fig.cap='Perceived meritocracy items'}
 load(here::here("input/data/proc/elsoc_w03.RData"))   
 pacman::p_load(dplyr,sjPlot,ggplot2)

  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,
           inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit() 
 dat_likert_s1 <-
   dplyr::select(elsoc_w03,meffort,mtalent)
 dat_likert_s1 <- sjmisc::rec(dat_likert_s1, rec="rev", append=FALSE)

@@

```{r corr-mat,  fig.cap='Correlation matrix between social status and perceived meritocracy', fig.height=8, fig.width=10, out.width='80%'}
 load(here::here("input/data/proc/elsoc_w03.RData"))
 pacman::p_load(dplyr,sjPlot,ggplot2,psych,polycor)
 # elsoc_w03$merit <- 
 #   (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2
  df_corr<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,ess,
               inc10h.imp,edcine2,egpisko05,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit() %>% 
    dplyr::select(
                  # -meffort,-mtalent,
                  -sexo,-edad,
                  -ppolcat,-lngap_perc
                  ) 
 cormat<- polycor::hetcor(data = df_corr,parallel = T)
 cormat <- as.matrix(cormat)

  diag(cormat) = NA #set diagonal values to NA
  # Set Row names of the matrix
  rownames(cormat) <- c("A. People are rewarded for their efforts",
                       "B. People are rewarded for their intelligence",
                       "C. Subjective social status",
                       "D. Household income decile", 
                       "E. Educational level",
                       "F. Social Class")
  #set Column names of the matrix
  colnames(cormat) <-c("(A)", "(B)","(C)","(D)","(E)","(F)")
  #Plot the matrix using corrplot
  corrplot::corrplot(cormat,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
  text(4.7, 0.6, paste0("Source: Authors calculation based on ELSOC 2018",
                             " (N = ",dim(df_corr)[1],")"),   
       cex = 0.75)

@@

```{r ols-merit, echo=FALSE, results='asis'}
pacman::p_load(ordinal,texreg,dplyr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit()

 or01tal  <- lm(merit ~inc10h.imp+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or02tal  <- lm(merit ~edcine2+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or03tal  <- lm(merit ~egpisko05+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or04tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or05tal  <- lm(merit ~ess+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or08tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or09tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+inc10h.imp*ess+edad+sexo+ppolcat+lngap_perc, data= elsoc_w03)
 or10tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+edcine2*ess+edad+sexo+ppolcat+lngap_perc,data= elsoc_w03)
 or11tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+egpisko05*ess+edad+sexo+ppolcat+lngap_perc,  data= elsoc_w03)
 or12tal  <- lm(merit ~inc10h.imp*ess+edcine2*ess+egpisko05*ess+ess+edad+sexo+ppolcat+lngap_perc,  data= elsoc_w03)

 # knitreg(l = list(or01tal,or02tal,or03tal,or05tal,or08tal,or09tal,or10tal,or11tal))

 coef.names02 <-c("Subjective Social Status (SSS) ",
                  "Household Income Decile",
                  "Primary and Lower secondary",
                  "Upper secondary",
                  "Short-cycle tertiary",
                  "Tertiary or higher",
                  "Skilled Manual Worker (V+VI)",
                  "Self-Employment (IV.ab)",
                  "Routine Non-manual (III)",
                  "Service (I+II)",
                  "Int. (Income Decile x Subj. Status.)",
                  "Primary and Lower secondary x SSS",
                  "Upper secondary x SSS",
                  "Short-cycle tertiary x SSS",
                  "Tertiary or higher x SSS",
                  "Skilled Manual Worker (V+VI) x SSS",
                  "Self-Employment (IVab) x SSS",
                  "Routine Non-manual (III) x SSS",
                  "Service (I+II) x SSS")

 tab.caption02 <- "Linear Regression for Perception of Meritocracy"
 omit.coef02 <- c("((Intercept))|(edad)|(sexo)|(ppolcat)|(ppolcat)|(lngap_perc)|(estlab)")

 knitreg(l = list(or05tal,or01tal,or02tal,or03tal,or08tal,or09tal,or10tal,or11tal),
        include.thresholds = FALSE,
        custom.coef.names = coef.names02,
        caption.above = TRUE,
        caption = paste("(\\#tab:merit)",tab.caption02),
        float.pos = "h!",
        custom.model.names = paste0("Model ",c(1:8)),
        omit.coef = omit.coef02,
        scalebox = 0.7,
        groups = list("Education (ref: Incomplete Primary or lower)" = 3:6,
                      "Social Class (ref: Unskilled worker)" =7:10,
                      "Int. (Education x Subj. Status.)" =12:15,
                      "Int. (Class x Subj. Status.)" =16:19),
        custom.gof.rows = list("Controls"=rep("Yes",8))
        )

@@

```{r int-income, fig.cap='Conditional effects of income decile by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
pacman::p_load(ordinal,texreg,interplot,margins,dplyr,sjPlot) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit()

 or09tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+inc10h.imp*ess+edad+sexo+ppolcat+lngap_perc, 
                data= elsoc_w03)
 # Income......................................................................
 marg09tal<- interplot(m = or09tal, var1 = "inc10h.imp", var2 = "ess",
                       point = T,esize = 1,hist = F) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   scale_x_continuous(name = NULL,
                      breaks = 0:10,labels = c("Bottom", 1:9,"Top")) +
   # ylab("A")+
   theme(axis.text=element_text(size=18),
         axis.title.y=element_text(size = 20,face = "bold", angle=0,vjust = 1))

 pred09tal<-
   plot_model(model = or09tal,
              type = "int",
              legend.title = "Subj. Status") +
   labs(title =NULL) + ylab(label = "Perceived Meritocracy") +
   ylim(1, 5) + 
   geom_point() +
   scale_x_continuous(name = "Income Decile",
                      breaks = 1:10,labels = paste0("D",1:10)) +
   scale_color_discrete(guide = guide_legend(title.position = "top",
                                             direction = "horizontal",
                                             title.hjust = 0.5)) +
   theme(legend.position=c(0.1, 0.1),
         axis.text=element_text(size=15),
         legend.background = element_blank(),
         legend.box.background = element_rect(colour = "black"))

  cowplot::plot_grid(
   # marg09tal,
    pred09tal,
                  # labels=c("A"," ","B"," ","C"," "),
                  nrow = 1,label_size = 25) 
```

@@

<!-- In this sense, a relevant result is that the relation of income becomes positive (negative) and statistically significant for those who perceive themselves as above (below) the median (_Mdn_ = 5; _M_ = 4.5) -->

 <!-- The linear prediction on the right column of Panel A shows the moderating effect of subjective social status on income, where those to belong to the richer income deciles and perceive themselves to be at the top perceive higher meritocracy compared to those as affluent but see themselves at the bottom. -->

<!-- Esto significa que para aquellos que perciben bajo estatus el efecto del ingreso se va volviendo más negativo en percepción de meritocracia, y serían aquellos que también serían "de altos ingresos" y demandarían mayor redistribución

En la misma línea, aquellos que perciben alto estatus, en la medida que aumenta el ingreso la percepción de meritocracia va siendo mucho más alta. Los predicted values mostraban que  la diferencia principalmente se daba desde el Decil 6 en adelante.
-->
Model 7 shows that subjective social status positively moderates the negative influence of educational level on perceived meritocracy, particularly among tertiary education groups. Figure \@ref(fig:int-educ) shows slight differences in perceived meritocracy among groups with secondary or less education based on subjective social status. However, for short-cycle and tertiary education groups, perceived meritocracy is higher for those with high subjective social status and lower for those with low subjective social status. The negative association of education can be observed among individuals with tertiary education who perceive themselves as close to the bottom of society and have a critical view of how merit is rewarded.

<!-- In other words, the degree of perceived meritocracy among lower-educated individuals is not affected by subjective status but is among those with higher educational attainment.

The results in Figure \@ref(fig:int-educ) depict that the moderating role of subjective status is more relevant among those who perceive themselves on the lower (0-2) and middle rungs (3-5), being more prominent among tertiary education groups, but still present in those with highschool education or lower. -->

```{r int-educ, fig.cap='Conditional effects of Education by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
pacman::p_load(ordinal,texreg,interplot,margins,dplyr,sjPlot) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit()

 or10tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+edcine2*ess+edad+sexo+ppolcat+lngap_perc,
                data= elsoc_w03)
# Education...................................................................
 lbeduc <- c("Incomplete Primary\n  or  less","Primary &\n Lower secondary",
             "Upper\n secondary ","Short-cycle\n tertiary","Tertiary\n or higher")

 marg10tal<- 
   interplot(m = or10tal, var1 = "edcine2", var2 = "ess", point = T,
             facet_labs = c("Primary &\n Lower secondary",
                            "Upper\n secondary ",
                            "Short-cycle\n tertiary","Tertiary\n or higher")
             ) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   # ylab("B")+    
   scale_x_continuous(name = NULL,
                      breaks = 0:10,labels = c("B", 1:9,"T")) +
   theme(strip.text.x = element_text(size = 17),
         axis.text=element_text(size=17),
         axis.title.y=element_text(size = 20,face = "bold", angle=0,vjust = 1.2))

 pred10tal<- plot_model(model = or10tal,
                        type = "int",
                        legend.title = "Subj. Status") +
  labs(title =NULL) +  
  geom_line() +
  ylab(label = "Perceived Meritocracy")+
  ylim(1, 5) +   
  scale_x_discrete(name=NULL,limits=1:5, labels = lbeduc) +
  scale_color_discrete(guide = guide_legend(title.position = "top",
                                            direction = "horizontal",
                                            title.hjust = 0.5)) +
  theme(
    legend.position=c(0.1, 0.1),
    axis.text=element_text(size=15),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black"))

   cowplot::plot_grid(
    #marg10tal,
    pred10tal,
                  # labels=c("A"," ","B"," ","C"," "),
                  nrow = 1,label_size = 25) 
```

@@

Subjective social status moderates the relationship between perceived meritocracy and routine non-manual and service classes, as shown by Model 8. Figure \@ref(fig:int-class) demonstrates that workers in these classes perceive lower meritocracy when they perceive themselves as lower on the social ladder and higher meritocracy when they perceive themselves as higher on the social ladder. Our findings did not show significant differences in perceived meritocracy across social classes. However, subjective social status does affect perceptions of meritocracy among clerks, sales workers, and higher-grade professionals, even when controlling for factors such as occupation, autonomy, and authority in the workplace.

```{r int-class, fig.cap='Conditional effects of Social class by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
pacman::p_load(ordinal,texreg,interplot,margins,dplyr,sjPlot) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit()

 or11tal  <- lm(merit ~inc10h.imp+edcine2+egpisko05+ess+egpisko05*ess+edad+sexo+ppolcat+lngap_perc, 
                data= elsoc_w03)

 #Social Class..................................................................
  labsegp <- c("Unskilled\n(VII.ab+IVc)",
              "Skilled Manual\n(V+VI)",
              "Self-Employment\n(IVab)",
              "Routine\nNon-manual (III)",
              "Service\n(I+II)")

>>>>

---
title: "Subjective status, Social Class and meritocracy perception"
linestretch: '1.5'          # interlineado
link-citations: yes         # citas linkeadas
fontsize: 12pt
linkcolor: blue
author:
- name: Julio Iturra ^[Corresponding author - jiturra@bigsss.uni-bremen.de]
  affiliation: Bremen International Graduate School of Social Sciences
- name: Juan Carlos Castillo
  affiliation: Universidad de Chile
bibliography: "input/bib/ess-merit.bib"
csl: "input/bib/apa6.csl"
css: "input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::word_document2:
    reference_docx: input/template.docx
    toc: false
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: false
    pandoc_args:
    - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
geometry: margin=1in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}  
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
lang: en-US
abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
keywords: subjective social status, meritocracy perception, social class
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

@@

@@

The objective position is measured through three variables. First, _Household Income Decile_ is calculated using monthly per capita household income and categorised into deciles. Second, _Educational level_ corresponds to the five categories for educational attainment following the International Standard Classification of Education [@unesco_international_2012].

Control variables are age and gender

@@

```{r plot-likert, fig.height=3, fig.width=10, fig.cap='Perceived meritocracy items'}
 load(here::here("input/data/proc/elsoc_w03.RData"))   
 pacman::p_load(dplyr,sjPlot,ggplot2)
 elsoc_w03$Q05h[elsoc_w03$Q05h=="No responde"] <- NA
  elsoc_w03$merit <-(as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,merit,income=inc10h, edcine2, ess, sexo, edad,ponderador02) %>% 
        na.omit() 
 dat_likert_s1 <-
   dplyr::select(elsoc_w03,meffort,mtalent)
 dat_likert_s1 <- sjmisc::rec(dat_likert_s1, rec="rev", append=FALSE)

@@

```{r corr-mat,  fig.cap='Correlation matrix between social status and perceived meritocracy', fig.height=8, fig.width=10, out.width='80%'}
 load(here::here("input/data/proc/elsoc_w03.RData"))
 pacman::p_load(dplyr,sjPlot,ggplot2,psych,polycor)
 elsoc_w03$merit <-(as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2
  df_corr<- elsoc_w03 %>%
    dplyr::select(merit,income=inc10h, edcine2,ess, sexo, edad) %>% 
        na.omit() %>% 
    dplyr::select(-sexo,-edad) 
 cormat<- polycor::hetcor(data = df_corr,parallel = T)
 cormat <- as.matrix(cormat)

  diag(cormat) = NA #set diagonal values to NA
  # Set Row names of the matrix
  rownames(cormat) <- c("A. Meritocracy perception",
                        "B. Household income decile",
                        "C. Educational level",                        
                        "D. Subjective social status"
                        )
  #set Column names of the matrix
  colnames(cormat) <-c("(A)", "(B)","(C)","(D)")
  #Plot the matrix using corrplot
  corrplot::corrplot(cormat,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
  text(1, 0.5, paste0("Source: Authors calculation based on ELSOC 2018",
                             " (N = ",dim(df_corr)[1],")"),   
       cex = 1)

@@

```{r ols-merit, echo=FALSE, results='asis'}
pacman::p_load(texreg,dplyr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess, sexo, edad,ponderador02) %>%
  na.omit()

dfreg$income[dfreg$income=="99"] <- NA
dfreg$ess_cwc <- mean(dfreg$ess) - dfreg$ess

m_inc <- lm(merit~income,data = dfreg,weights =ponderador02)
m_educ <- lm(merit~edcine2,data = dfreg,weights =ponderador02)
m_status <- lm(merit~income+edcine2+ess+sexo+edad,data = dfreg,weights =ponderador02) 
m_inter_inc <- lm(merit~income*ess+edcine2+ess+sexo+edad,data = dfreg,weights =ponderador02)
m_inter_edu <-lm(merit~income+edcine2*ess+ess+sexo+edad,data = dfreg,weights =ponderador02) 
models <- list(m_inc,m_educ,m_status,m_inter_inc,m_inter_edu)
caption <- "Linear Regression for Perception of Meritocracy"

knitreg(
  l = models,
  caption = paste("(\\#tab:merit)",caption),
  caption.above = T,
  custom.coef.map = list(
    income = "Income Decile",
    edcine2cine12 ="Primary & Lower secondary",
    edcine2cine34 = "Upper secondary",
    edcine2cine5 = "Short-cycle tertiary",
    edcine2cine678 = "Tertiary or higher",
    ess="Subjective Social Status",
    "income:ess" = "Income x SSS",
    "edcine2cine12:ess"= "Primary & Lower secondary x SSS",
    "edcine2cine34:ess"= "Upper secondary x SSS",
    "edcine2cine5:ess"= "Short-cycle tertiary x SSS",
    "edcine2cine678:ess"= "Tertiary or higher x SSS"
  ),
  groups = list("Education (ref: Incomplete Primary or less)"=2:5)
)
```

@@

```{r int-income, fig.cap='Conditional effects of income decile by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(theme_minimal())
ggplot2::theme_update(text=element_text(size=15,  family="serif")) 
pacman::p_load(texreg,dplyr,marginaleffects,ggplot2) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2

elsoc_w03$ess_cwc <- sjmisc::center(x = elsoc_w03$ess)
summary(elsoc_w03$ess_cwc)

dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess=ess_cwc, sexo, edad,ponderador02) %>%
  na.omit()
ess_max<- mean(dfreg$ess) + sd(dfreg$ess)
ess_min<- mean(dfreg$ess) - sd(dfreg$ess)
ess_mea<- mean(dfreg$ess)

dfreg$income[dfreg$income=="99"] <- NA
m_inter_inc <- lm(merit~income*ess+edcine2+ess+sexo+edad,data = dfreg,weights=ponderador02)

df_pref <-
  predictions(
    m_inter_inc,newdata = datagrid(
      ess = c(ess_min, ess_mea, ess_max),
      income = 1:10,
      edcine2 = "cine34",
      sexo = "Female",
      edad = mean(dfreg$edad)
    )
  )

df_pref$ess <- as.factor(df_pref$ess)
df_pref %>% 
ggplot(aes(y=estimate,x=income, fill=ess,color=ess,ymin=conf.low, ymax=conf.high)) +
  geom_line(aes(y=estimate,x=income),size=0.75) +
  geom_point(size=2) +
  geom_ribbon(alpha=0.2,size=0.5,linetype=1) +
  scale_color_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD")) +
  scale_fill_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD"))+
   labs(y = "Perceived Meritocracy",x="Income Decile",) +
   ylim(1, 5) +
   scale_x_continuous(breaks = 1:10,labels = paste0("D",1:10)) +
   theme(legend.position="right",
         legend.direction = "vertical",
         axis.text=element_text(size=15),
         legend.background = element_blank(),
         legend.box.background = element_rect(colour = "black"))
```

@@

Model 7 shows that subjective social status positively moderates the negative influence of educational level on perceived meritocracy, particularly among tertiary education groups. Figure \@ref(fig:int-educ) shows slight differences in perceived meritocracy among groups with secondary or less education based on subjective social status. However, for short-cycle and tertiary education groups, perceived meritocracy is higher for those with high subjective social status and lower for those with low subjective social status. The negative association of education can be observed among individuals with tertiary education who perceive themselves as close to the bottom of society and have a critical view of how merit is rewarded.

```{r int-educ, fig.cap='Conditional effects of Education by subjective social status on perceived meritocracy', fig.width=10, fig.height=5}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(theme_minimal())
ggplot2::theme_update(text=element_text(size=15,  family="serif")) 
pacman::p_load(texreg,dplyr,marginaleffects,ggplot2,stringr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
elsoc_w03$ess_cwc <- sjmisc::center(x = elsoc_w03$ess)

dfreg <- elsoc_w03 %>%
  dplyr::select(merit,income=inc10h, edcine2, ess=ess_cwc, sexo, edad,ponderador02) %>%
  na.omit()
dfreg$ess_cwc <- mean(dfreg$ess) - dfreg$ess
ess_max<- mean(dfreg$ess) + sd(dfreg$ess)
ess_min<- mean(dfreg$ess) - sd(dfreg$ess)
ess_mea<- mean(dfreg$ess)
dfreg$income[dfreg$income=="99"] <- NA
m_inter_edu <- lm(merit~income+edcine2*ess+ess+sexo+edad,data = dfreg,weights=ponderador02)

df_pref <-
  predictions(
    m_inter_edu,newdata = datagrid(
      ess = c(ess_min, ess_mea, ess_max),
      income = 6,
      edcine2 = levels(dfreg$edcine2),
      sexo = "Female",
      edad = mean(dfreg$edad)
    )
  )

df_pref <- as.data.frame(df_pref)
df_pref$ess <- as.factor(df_pref$ess)

df_pref$edcine2 <- factor(df_pref$edcine2,levels = ,labels = c("Incomplete Primary or less","Primary & Lower secondary","Upper secondary ","Short-cycle tertiary","Tertiary or higher"))

df_pref %>% 
ggplot(aes(y=estimate,x=edcine2, color=ess, group=ess,ymin=conf.low, ymax=conf.high)) +
  geom_line(size=0.75) +
  geom_point(size=2) +
  geom_errorbar(alpha=1,size=1,width=.1,linetype=1) +
  scale_color_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD")) +
  scale_fill_discrete(name="Sub.Status",labels = c("-1SD","Mean","+1SD"))+
  labs(y = "Perceived Meritocracy",x="Education") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
     theme(legend.position="right",
         legend.direction = "vertical",
         axis.text=element_text(size=15),
         legend.background = element_blank(),
         legend.box.background = element_rect(colour = "black"))
```

@@
